Variance reduction for MCMC methods via martingale representations

### 2020-03-25 aap/Revision/vr-mcmc-revision-new-new.tex

\frac{\mathrm{cost}(\pi_{K,n}^N) \PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]}{\mathrm{cost}(\pi_{n}^N)\PVar[\pi_{n}^N(f)]}\leq C e^{\varkappa n\gamma^2} K \gamma \, T^2 \mathtt{D}_\Psi. 

@@

Thus, for any fixed $K\geq 1,$  our variance reduction method is advantageous if $\gamma<\min\{1/(\mathtt{D}_\Psi T^2),1/\sqrt{n}\}.$  

@@

\subsection{Banana shape distribution}
The “Banana-shape” distribution, proposed by \cite{Haario1999}, can be obtained from a d-dimensional Gaussian vector with zero mean and covariance $\mathrm{diag}(p,1,\ldots,1)$ by applying transformation $\varphi_b(x) : \rset^d \rightarrow \rset^d$ of the form
\[
\varphi(x_1,\ldots,x_n) = (x_1, x_2 + bx^2_1 - pb,x_3,\ldots,x_d)
\]
where $p > 0$ and $b > 0$ are parameters; $b$ accounts for the curvature of density’s level sets. The potential $U$ is given by
\[
U(x_1, . . . , x_d) = \frac{x_1^2}{2} + (x_2 + bx_1^2 - pb)^2 + \sum\limits_{k=3}^{d}\frac{x_k^2}{2}
\]
The quantity of interest is the expectation of $f(x) = x_2$. We set $p = 100, b = 0.1$ and consider $d=8$. We solve the least squares  problems \eqref{eq:Qregr} using the second-order polynomial approximations for the coefficients \(\hat{a}_{p,\mathbf{k}}\) as described in the previous section, hence we set $K = 2$. We use $T=10$ independent training trajectories, each of size $n=10^4$. We set the truncation level \(n_{\mathrm{trunc}} = 50\). To test our variance reduction algorithm, we generated 100 independent trajectories of length $n = 10^3$ and plot boxplots of the ergodic averages in Figure~\ref{fig:banana_shape}.

>>>>

\frac{\mathrm{cost}(\widehat \pi_{K,n}^N) \PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]}{\mathrm{cost}(\pi_{n}^N)\PVar[\pi_{n}^N(f)]}\leq C e^{\varkappa n\gamma^2} K \gamma \, T^2 \mathtt{D}_\Psi. 

@@

Thus, for any fixed $K\geq 1,$  our variance reduction method is advantageous if $\gamma\ll\min\{1/(\mathtt{D}_\Psi T^2),1/\sqrt{n}\}.$  

@@

\subsection{Banana shape distribution}
The “Banana-shape” distribution, proposed by \cite{haario1999adaptive}, can be obtained from a d-dimensional Gaussian vector with zero mean and covariance $\mathrm{diag}(p,1,\ldots,1)$ by applying transformation $\varphi_b(x) : \rset^d \rightarrow \rset^d$ of the form
\[
\varphi(x_1,\ldots,x_n) = (x_1, x_2 + bx^2_1 - pb,x_3,\ldots,x_d)
\]
where $p > 0$ and $b > 0$ are parameters; $b$ accounts for the curvature of density’s level sets. The potential $U$ is given by
\[
U(x_1, . . . , x_d) = \frac{x_1^2}{2} + (x_2 + bx_1^2 - pb)^2 + \sum\limits_{k=3}^{d}\frac{x_k^2}{2}
\]
The quantity of interest is the expectation of $f(x) = x_2$. We set $p = 100, b = 0.1$ and consider $d=8$. We solve the least squares  problems \eqref{eq:Qregr} using the second-order polynomial approximations for the coefficients \(\hat{a}_{p,\mathbf{k}}\) as described in the previous section, hence we set $K = 2$. We use $T=10$ independent training trajectories, each of size $n=10^4$. We set the truncation level \(n_{\mathrm{trunc}} = 50\). To test our variance reduction algorithm, we generated 100 independent trajectories of length $n = 10^3$ and plot boxplots of the ergodic averages in Figure~\ref{fig:banana_shape}.
